\documentclass{article}
\usepackage{enumitem}
\usepackage{amsmath,amssymb,graphicx}
\usepackage{Sweave}
\usepackage{tikz}
\usetikzlibrary{positioning,shapes.geometric,decorations.text}
\begin{document}
%\input{Project-Proposal-concordance}

\newcommand{\E}{\mathbb{E}}
\newcommand{\indp}{\ensuremath{\perp\hspace{-5pt}\perp}}
\newcommand{\M}{\mathcal{M}}
\newcommand{\given}{\; \mid \;}



\title{\textbf{The Effect of Dieting on Sleep}}
\author{Alex Luedtke, Lucia Petito, and Steven Pollack}
\date{}
\maketitle

\section{Specify the Question}

We know that sleep habits affect weight, but does trying to lose weight affect sleep patterns?

Sleep and food consumption are two things every person thinks about each day, usually with the attitude ``I should eat less and sleep more.''  Our interest stems from a lack of research into this question.  Researchers have conducted studies in the reverse direction, trying to relate sleeping habits to successful dieting.  A relationship between obesity and sleep cycles has been established: both oversleeping and sleep deprivation are associated with an increased Body Mass Index (BMI) [REFERENCE].  

There are reasonable explanations for all possible outcomes of this analysis.  An increased interest in one's diet could indicate an increased interest in one's overall health.  Because doctors advise their patients to consistently get enough sleep each night to maintain a healthy lifestyle, it would be reasonable to see more hours of sleep in people who are dieting.  On the other hand, dieting has profound psychological and hormonal effects, which could cause a disruption to normal sleep patterns and result in less sleep on average.

For the purposes of this analysis, our population of interest is adults ages 25-79 who are not pregnant, have never been diagnosed by a doctor with a sleep disorder, and live in the USA.  We used the NHANES data from 2009-2010 [1, 2] from the following questionnaires: ``Demographic Variables and Sample Weights," ``Weight History 16 Years and Older," and ``Sleep Disorders."

NHANES is a multi-stage survey of the U.S. population that the Center for Disease Control and Prevention (CDC) conducts every 2 years.  They sample counties first, then ``segments'' (which can be as small as street blocks to as big as neighborhoods: determined by population density), then households, and finally individuals.  It aims to study a wide range of topics such as cardiovascular disease, obesity, physical fitness and physical functioning, reproductive history and sexual behavior, etc.  Each individual participates in an interview and a physical examination.  The sample is chosen to represent the U.S. population of all ages.  To produce reliable statistics, NHANES over-samples persons 60 and older, adolescents, African Americans, and Hispanics among others.  

However, to protect the privacy of the participants, the CDC strips the data of any information that would allow an analyst to find an individual.  This process mostly removes geographic identifiers; in particular, they remove all information about the first three stages of their sampling.  Unfortunately, this removal of information means that we cannot replicate their sampling method accurately, which....

Our baseline covariates are:
\begin{itemize}
\item When the interview was conducted (1 for November 1, 2009 - April 30, 2010, 2 for May 1, 2009 - October 31, 2010)
\item Subject Gender (1 for Male, 2 for Female)
\item Age in Months (300-959, or 25-79 years)
\item Race/Ethnicity (1 for Mexican American, 2 for Other Hispanic, 3 for Non-Hispanic White, 4 for Non-Hispanic Black, 5 for Other Race)
\item Education Level (1: less than 9th grade, 2: 9-11th grade (including 12th grade with no diploma), 3: high school grad/GED or equivalent, 4: some college or AA degree, 5: college graduate or above)
\item Marital Status (1: married, 2: widowed, 3: divorced, 4: separated, 5: never married, 6: living with partner)
\item Annual Household Income (14 levels, \$0-\$24,999 by \$5,000, \$25,000-\$74,999 by \$10,000,  \$75,000 - \$99,999, \$100,000 and over, <\$20,000, >\$20,000)
\item Body Mass Index (continuous values from ~15-50) note: this is a combination of subject weight and subject height
\end{itemize}

This data set contains many cases that have missing data.  The percentage missing in each covariate is in the table below.  We decided not to attempt multiple imputation; we will only consider complete cases in this analysis.  Specifically, we deleted all people who responded ``refused" or ``don't know" or are missing data for Annual Household Income, Marital Status, Education Level, and Interview Time Period.  We deleted all people who were younger than 25 and older than 79, all pregnant women, and all people who had ever been diagnosed by a doctor with a sleep disorder (or who responded that they ``don't know").  

We also coarsened the categories for several of the variables.  We decided that we were interested in the ``Marital Status'' variable because it described whether or not the individual had a stable relationship in their life, so we collapsed it into ``married or living with partner'' vs ''else.''  Because a significant number of people replied ''<\$20,000'' and ''>\$20,000'' to the annual household income question, we had to define all individuals as such.  We also collapsed the bottom two categories for ''Education Level'' into one category: ''less than 11th grade.''

[INSERT TABLE]

% We have information about how much weight each person lost in the past year and whether or not that person lost the weight intentionally.  In our analysis we will treat whether or not a person tried to diet as the intervention variable. For all of our analyses, we will use the same outcome variable: average amount of sleep per night. However, we also have data on how much weight each individual lost in the past year (through the questions ``How much did you weigh one year ago'' and   ``How much do you weigh today'').  To capture how seriously someone is dieting, we will then perform a secondary analysis in which we discretize the amount of weight a person lost.  This will be a mediator variable, as it will fall between our intervention variable and our outcome variable.

% \section{Potential Sources of Bias}
% 
% We would have liked to exclude all people who are on medications that cause excessive weight gain as a side effect, such as anxiety and depression medications.  Unfortunately, our data set does not include this covariate.  
% 
% Our annual household income data 

\section{Specify the Causal Model}

Our observational data structure is $O=(W,A,Y) \sim P_0$. Our endogenous variables are $W$, $A$, and $Y$, where $W$ has the following dimensions:

\begin{itemize}
\item When the interview was conducted (1: November 1, 2009 - April 30, 2010, 2: May 1, 2009 - October 31, 2010)
\item Subject Gender (1: Male, 2: Female)
\item Age in Months (300-959, or 25-79 years)
\item Race/Ethnicity (1: Mexican American, 2: Other Hispanic, 3: Non-Hispanic White, 4: Non-Hispanic Black, 5: Other Race)
\item Education Level (1: less than 11th grade (including 12th grade with no diploma), 3: high school grad/GED or equivalent, 4: some college or AA degree, 5: college graduate or above)
\item Marital Status (1: married or living with partner, 0: else)
\item Annual Household Income (0: <\$20,000, 1: >\$20,000)
\item Body Mass Index (continuous values from ~15-50) [Note: this is a combination of subject weight and subject height]
\end{itemize}

$A$ is the individual's response to the survey question: ``During the past 12 months, have you tried to lose weight?", and $Y$ is the individual's response to ``How much sleep do you usually get at night on weekdays or workdays?''  

Our exogenous variables are $U = (U_W, U_A, U_Y) \sim P_U$ (we make no assumption on the distribution).  

We make no exclusion restrictions on our data and no assumptions on the functional forms of $W$, $A$, or $Y$, so with mild temporal assumptions, one SCM is:

\begin{align*}
W &= f_{W}(U_{W}) \\
A &= f_{A}(W,U_{A}) \\
Y &= f_{Y}(W,A,U_{Y}) \\
\end{align*}

\begin{figure}[h]
    \centering
    \begin{tikzpicture}[->, shorten >=2pt, >=stealth, node distance=1cm,
                        pil/.style={->,thick,shorten =2pt,},scale=0.65
                        ]
    \node[circle,draw] at (4,2) (1) {W};
    \node[circle,draw] at (2,0) (2) {A};
    \node[circle,draw] at (6,0) (3) {Y};
    \node[circle,draw] at (4,4) (4) {U};
    \draw[->] (1) to [out=195,in=95] (2);
    \draw[->] (2.east) -- (3.west);
    \draw[->] (1) to [out=345,in=95] (3);
    \draw[->] (4.south) -- (1.north);
    \draw[->] (4) to [out=210,in=165] (2);
    \draw[->] (4) to [out=330,in=30] (3);
    \end{tikzpicture}
  \caption{Simplified DAG -- no independence assumptions on $U$'s.}
  \label{fig:DAG}
\end{figure}


\section{Specify the Causal Parameter of Interest}

Since the intervention is a point treatment, our counterfactual is $Y_{a}$: the average sleep an individual would get, if they, possibly contrary to fact, had (or had not) attempted to lose weight.

We are interested in measuring the average treatment effect (ATE) of attempted weight loss on average sleep:

$$\Psi(P_{U,X}) = \E_{U,X}[Y_1] - \E_{U,X}[Y_0]$$

\section{Assess Identifiability and Positivity}

Since $\Psi$ is the ATE, we need to satisfy backdoor criterion to identify g-computation with $\Psi$:
   $$Y_{a} \indp A \mid W$$

For identfiability, we need $U_{A} \indp U_{Y}$ and either
    \begin{itemize}
      \item $U_{A} \indp U_{W}$  (semi-plausible)
      \item $U_{W} \indp U_{Y}$  (less plausible)
    \end{itemize}

\begin{center}
    \begin{tikzpicture}[->, shorten >=2pt, >=stealth, node distance=1cm,
                        pil/.style={->,thick,shorten =2pt,},scale=0.65,font=\scriptsize
                        ]
    \node (A) [circle,draw] {A};
    \node (W) [circle,draw,above right=of A] {W};
    \node (Y) [circle,draw,below right=of W] {Y};
    \node (UA) [circle,draw,below left=of A] {$U_A$};
    \node (UW) [circle,draw,above=of W] {$U_W$};
    \node (UY) [circle,draw,below right=of Y] {$U_Y$};
    \draw[->] (W) to [out=195,in=95] (A);
    \draw[->] (A.east) -- (Y.west);
    \draw[->] (W) to [out=345,in=95] (Y);
    \draw[->] (UW.south) -- (W.north);
    \draw[->] (UA) to [out=45,in=225] (A);
    \draw[->] (UY) to [out=135,in=315] (Y);
    \draw[dashed,<->] (UW) to [out=330,in=95] (UY);
    %\pgftransformyshift{-.65cm}
    %\draw[decoration={text along path, text={X},text align={center}},decorate] (UA) to
                      [out=345,in=195] (UY);
    \draw[dashed,<->,postaction={decorate,decoration={raise=-4pt,text along path,text align=center,text={X}}}] (UA) to [out=345,in=195] (UY);
    \draw[dashed,<->,postaction={decorate,decoration={raise=-4pt,reverse path,text along path,text align=center,text={???}}}] (UW) to [out=210,in=85] (UA);
    \draw[dashed,<->,
    postaction={decorate,
    decoration={raise=-4pt,text along path,text align=center,text={???}}}] (UW) to [out=330,in=95] (UY);
    \end{tikzpicture}
\end{center}

We must also address the issue of positivity.  We do not believe that trying to lose weight or not trying to lose weight occurs with probability $0$ given any set of covariates $W$.  Thus we believe our theoretical positivity requirements are met.  However, practical positivity violations are very prevalent given that $W$ is high dimensional.  If we only consider those variables in $W$ that are categorical (Gender, Race/Ethnicity, Education Level, Marital Status), we only have two strata where no people tried to lose weight:

\begin{tabular}{| l | l | l | l |}
\hline
 Gender & Race & Education & Marital Status \\
\hline
Male & Other & Less than HS & Unmarried/not living with partner \\
Female & Other & HS & Married/living with partner \\
\hline
\end{tabular}

\section{Commit to a Statistical Model and Target Parameter of the Observed Data Distribution}

We still have no assumptions on the functional forms of $f_W, f_A$, and $f_Y$, so we choose the non-parametric model $\mathcal{M}$ of distributions compatible with our SCM.  

We are still interested in the ATE, so our target parameter of the observed data distribution (assuming that we have a simple random sample of 4,077 individuals) is

$$\psi_0 = \E_{X,U}[ \E$$

\section{Estimate the Chosen Parameter of the Observed Data Distribution}

Provided we can accept $U_A \indp U_W$ or $U_W \indp U_Y$, we will estimate the ATE in the following ways: 

\begin{itemize}
  \item Simple Substitution:
    \[
      \psi_0 \approx \frac{1}{n}\sum_{i=1}^{n}\widehat{\E}\left[Y \mid A=1, W=W_i \right] - \widehat{\E}\left[Y \mid A=0, W=W_i\right]
    \]
  \item IPTW:
  \[
    \psi_{0} \approx \frac{1}{n}\sum_{i=1}^{n} \left(\frac{I(A_i=1)}{\hat{g}_n(A_i \mid W_i)} - \frac{I(A_i=0)}{\hat{g}_n(A_i \mid W_i)} \right)Y_i
  \]
  \item TMLE:
  \[
    \psi_{0} \approx \frac{1}{n}\sum_{i=1}^{n}\left( \bar{Q}_{n}^{1}(1,W_i) - \bar{Q}_{n}^{1}(0,W_i)\right)
  \]
\end{itemize}

%   where
%   \[
%   \bar{Q}_{n}^{1}(A,W) = \logit(\bar{Q}_{n}^{0}(A,W)) + \epsilon_{n}\left(\frac{I(A=1)}{\hat{g}_n(A=1\mid W)} - \frac{I(A=0)}{\hat{g}_n(A=0\mid W)} \right)
%   \]

\section{Interpret Results}




%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Indicators that a person is trying to diet}

Below are the potential responses to the question ``How did you try to lose weight?" [2]:

\begin{itemize}[noitemsep]
\item Ate less food
\item Switched to foods with lower calories
\item Ate less fat
\item Ate fewer carbohydrates
\item Exercised
\item Skipped meals
\item Ate ``diet'' foods or products
\item Used a liquid diet formula such as slimfast or optifast
\item Joined a weight loss program such as Weight Watchers, Jenny Craig, Tops, or Overeaters Anonymous
\item Followed a special diet such as Dr. Atkins, South Beach, other high protein or low carbohydrate diet, cabbage soup diet, Ornish, Nutrisystem, Body-for-Life
\item Took diet pills prescribed by a doctor
\item Took other pills, medicines, herbs, or supplements not needing a prescription
\item Started to smoke or began to smoke again
\item Took laxatives or vomited
\item Drank a lot of water
\item Ate more fruits, vegetables, salads
\item Ate less sugar, candy, sweets
\item Changed eating habits (didn't eat late at night, ate several small meals a day)
\item Ate less junk food or fast food
\item Other (specify)
\end{itemize}

\subsection*{}
\begin{enumerate}
	\item Centers for Disease Control and Prevention (CDC). National Center for Health Statistics (NCHS). National Health and Nutrition Examination Survey Data. Hyattsville, MD: U.S. Department of Health and Human Services, Centers for Disease Control and Prevention, 2009-2010. \\ http://www.cdc.gov/nchs/nhanes/nhanes2009-2010/nhanes09\_10.htm.
	\item Centers for Disease Control and Prevention (CDC). National Center for Health Statistics (NCHS). National Health and Nutrition Examination Survey Questionnaire (or Examination Protocol, or Laboratory Protocol). Hyattsville, MD: U.S. Department of Health and Human Services, Centers for Disease Control and Prevention, 2009-2010. \\ http://www.cdc.gov/nchs/nhanes/nhanes2009-2010/questexam09\_10.htm.
\end{enumerate}

\end{document}
